---
title: "TFM"
author: "Maria del Mar Escalas Martorell"
date: "`r Sys.Date()`"
output: html_document
description: "This is the second .Rmd document from a set of three .Rdm. Code provided is complemetary to the main document containing the Master's Thesis in *Forecasting dwelling prices in the City of Madrid: a comprehensive analysis of the Real Estate market*, which can be found in this repository. This markdown contains the code needed for conducting the analysis of point 3. DATA EXTRACTION AND PREPROCESSING."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.align = "center")
```

Load libraries.

```{r message = FALSE, warning = FALSE}
# General libraries:
library(tidyverse)
library(stringr) # regex
library(sf) # geometry polygons and plots (maps)

# Dates
library(zoo)

# Web scraping libraries:
library(xml2)
library(httr)
library(scrapex)
library(rvest) # read_html

# Plots and maps
library(leaflet)

# Times New Roman font import:
library(extrafont)
```


## 3. DATA EXTRACTION AND PREPROCESSING

### 3.1. Dwellings' characteristics data

#### 3.1.1. Idealista (2023)

Extraction of data from a private API: 

- Reading personal private credentials (hidden): user's personal credentials are provided by API's host. It is shown below how they should be loaded in R from a hidden location in the computer. In this case, credentials are stored in a separated folder (not included in the data available in the repository). Instead of using this directory, the person who wishes to reproduce this code should introduce its own credentials. 

```{r}
# consumer_apikey <- readLines("../Secret/consumer_apikey.txt")
# consumer_secret <- readLines("../Secret/consumer_secret.txt")
# api_key_secret <- paste0(consumer_apikey, ":", consumer_secret)
```

- Using basic authentication: credentials should be encoded and the token needs to be build.

```{r}
# secret <- base64_enc(api_key_secret)
# 
# req <- httr::POST("https://api.idealista.com/oauth/token",
#        httr::add_headers(
#          "Authorization" = paste("Basic", secret, sep = " "),
#          "Content-Type" = "application/x-www-form-urlencoded;charset=utf-8"
#             ),
#        body = "grant_type=client_credentials&scope=read"
#             ) 
# 
# token <- paste("Bearer", httr::content(req)$access_token)
```

- Start building the request: establishing the desired parameters to search.

```{r}
# site = "https://api.idealista.com/3.5/es/search?" # official site provided by the host
# loc = "center=" # location reference
# x = 40.40691 # latitude
# y = -3.70958 # longitude
# country = "&country=es" # Spain
# maxItems = "&maxItems=50" # maximum items allowed for the request to return = 50
# pages = "&numPage=1"
# dist = "&distance="
# distance = 1000 # distance from the location reference (radius)
# property = "&propertyType="
# type = "homes"
# operation = "&operation="
# op = "sale" # only sales (not rentals)
# chalet = "&chalet=0" (no chalets)
```

- Building the url.

```{r}
# url1 <- paste(site, loc, x, ',', y, country, maxItems, pages, dist, distance, 
#              property, type, operation, op, sep = "")
```

- Requesting the information through Idealista's API.

```{r}
# res1 <- httr::POST(url1, httr::add_headers("Authorization" = token))
```

- Extracting the JSON content (a dataframe is obtained).

```{r}
# cont_raw2 <- httr::content(res1)
```

#### 3.1.2. Idealista (2018)

Database from Idealista (private online Real Estate portal).

Load needed package.

```{r}
# devtools::install_github("paezha/idealista18")
library(idealista18)
```

Data from Madrid 2018.

```{r}
data("Madrid_Sale")
class(Madrid_Sale)
```

Initial selection: from 42 variables to 30.

```{r}
idealista18 <- Madrid_Sale %>% 
  select(ASSETID, PERIOD, PRICE, CONSTRUCTEDAREA, UNITPRICE, ROOMNUMBER, 
         BATHNUMBER, HASTERRACE, HASLIFT, HASAIRCONDITIONING,
         ISPARKINGSPACEINCLUDEDINPRICE, HASNORTHORIENTATION, 
         HASSOUTHORIENTATION, HASEASTORIENTATION, HASWESTORIENTATION, 
         HASBOXROOM, HASWARDROBE, HASSWIMMINGPOOL, HASDOORMAN, HASGARDEN, 
         FLATLOCATIONID, CADCONSTRUCTIONYEAR, CADMAXBUILDINGFLOOR, 
         CADDWELLINGCOUNT, DISTANCE_TO_CITY_CENTER, DISTANCE_TO_METRO, 
         LONGITUDE, LATITUDE) %>% 
  rename(id = ASSETID, period = PERIOD, price = PRICE,
         built_area = CONSTRUCTEDAREA, price_sq_m = UNITPRICE,
         n_room = ROOMNUMBER, n_bath = BATHNUMBER, terrace = HASTERRACE,
         lift = HASLIFT, air_cond = HASAIRCONDITIONING,
         parking = ISPARKINGSPACEINCLUDEDINPRICE,
         north_orientation = HASNORTHORIENTATION,
         south_orientation = HASSOUTHORIENTATION, 
         east_orientation = HASEASTORIENTATION,
         west_orientation = HASWESTORIENTATION, boxroom = HASBOXROOM,
         wardrobe = HASWARDROBE, pool = HASSWIMMINGPOOL, 
         doorman = HASDOORMAN, garden = HASGARDEN,
         external = FLATLOCATIONID, year_built = CADCONSTRUCTIONYEAR,
         floors = CADMAXBUILDINGFLOOR, n_dwelling = CADDWELLINGCOUNT,
         km_to_center = DISTANCE_TO_CITY_CENTER,
         metro_proximity = DISTANCE_TO_METRO, long = LONGITUDE, lat = LATITUDE)
```

Issues with raw data:

- There is a difference between unique row number and the total number of rows.

```{r}
nrow(idealista18) - length(unique(idealista18$id))
```

- There are many dwellings without orientation.

```{r}
idealista18 %>% filter(north_orientation == 0 & south_orientation == 0 & east_orientation == 0 & west_orientation == 0)
```

- Dropping geometry to make operations.

```{r}
idealista18 <- st_drop_geometry(idealista18)
```

- Recoding of external. If "yes" = 1, if "no" = 0.

```{r}
idealista18 <- idealista18 %>% 
  mutate(external = case_when(external == 2 ~ 0,
                              external == 1 ~ 1,
                              is.na(external) ~ NA_real_))
```

Operations:

- Max for logical variables (in numeric format)
- Average for numerical (continuous) variables

```{r}
idealista <- idealista18 %>% 
  group_by(id) %>% 
  mutate(across(c(period, price, built_area, n_room, n_bath, 
                  year_built, floors, n_dwelling, 
                  km_to_center, metro_proximity, long, lat), mean),
         across(c(terrace, lift, air_cond, parking, 
                  north_orientation, south_orientation, east_orientation, 
                  west_orientation, boxroom, wardrobe, pool, doorman, 
                  garden, external), ~max(.)),
         across(c(price, built_area, price_sq_m, n_room, n_bath), round))

sapply(idealista, function(x) sum(is.na(x))*100/nrow(idealista)) # to check % of NAs per row

idealista <- idealista %>% distinct(id, .keep_all = TRUE) %>% drop_na() # to drop rows with no information available about 'external'
```

- Classifying orientation:

```{r}
idealista <- idealista %>% 
  mutate(orientation = as_factor(case_when(
  north_orientation == "0" & south_orientation == "0" & east_orientation == "0" & west_orientation == "0" ~ "None",
  north_orientation == "1" & south_orientation == "0" & east_orientation == "0" & west_orientation == "0" ~ "Low", # north
  north_orientation == "0" & south_orientation == "0" & east_orientation == "0" & west_orientation == "1" ~ "Medium", # west
  north_orientation == "0" & south_orientation == "0" & east_orientation == "1" & west_orientation == "0" ~ "Medium", # east
  north_orientation == "0" & south_orientation == "0" & east_orientation == "1" & west_orientation == "1" ~ "Medium", # east-west
  north_orientation == "1" & south_orientation == "0" & east_orientation == "0" & west_orientation == "1" ~ "Medium", # north-west
  north_orientation == "1" & south_orientation == "0" & east_orientation == "1" & west_orientation == "0" ~ "Medium", # north-east
  north_orientation == "1" & south_orientation == "0" & east_orientation == "1" & west_orientation == "1" ~ "Medium", # north-east-west
  north_orientation == "0" & south_orientation == "1" & east_orientation == "0" & west_orientation == "0" ~ "Good", # south
  north_orientation == "0" & south_orientation == "1" & east_orientation == "0" & west_orientation == "1" ~ "Good", # south-west
  north_orientation == "0" & south_orientation == "1" & east_orientation == "1" & west_orientation == "0" ~ "Good", # south-east
  north_orientation == "1" & south_orientation == "1" & east_orientation == "0" & west_orientation == "0" ~ "Good", # south-north
  north_orientation == "0" & south_orientation == "1" & east_orientation == "1" & west_orientation == "1" ~ "Good", # south-east-west
  north_orientation == "1" & south_orientation == "1" & east_orientation == "1" & west_orientation == "0" ~ "Good", # south-north-east
  north_orientation == "1" & south_orientation == "1" & east_orientation == "0" & west_orientation == "1" ~ "Good", # south-north-west
  north_orientation == "1" & south_orientation == "1" & east_orientation == "1" & west_orientation == "1" ~ "Full")),
         period = case_when(period == 201803 ~ "20181",
                            period == 201806 ~ "20182",
                            period == 201809 ~ "20183",
                            period == 201812 ~ "20184"),
         period = as.yearqtr(format(period), "%Y%q")) %>% 
  select(-c(north_orientation, south_orientation, east_orientation, west_orientation))
```

- Reconverting the dataframe into a geometry object.

```{r}
idealista <- st_as_sf(idealista, coords = c("long", "lat"), crs = 4326)
```

- Joining each dwelling with its corresponding neighbourhood. 

```{r}
madrid_idealista <- st_join(idealista, Madrid_Polygons, join = st_within)

madrid_idealista <- madrid_idealista %>% 
  select(-c(LOCATIONID, ZONELEVELID)) %>% 
  rename(neighbourhood = LOCATIONNAME)
```

- Incorporating coordinates again.

```{r}
madrid_idealista <- madrid_idealista %>%
  mutate(long = st_coordinates(geometry)[, 1],
         lat = st_coordinates(geometry)[, 2])
```

Identification of observations that contain errors:

- A10991866817670907438: price = 450000, n_rooms = 93 (wrong number of rooms)
- A8866129769871540210: price = 109000, n_rooms = 33 (wrong number of rooms) 

```{r}
madrid_idealista %>% filter(n_room > 25)
```

- A7501251371917084834: long = - 2.753303, lat = 36.75639 (house located in Almería)

```{r}
madrid_idealista %>% filter(long > -3.50 & lat < 40.00)
```

- Removing wrong observations.

```{r}
madrid_idealista <- madrid_idealista %>%
  filter(!(long > -3.50 & lat < 40.00)) %>%
  filter(!(price %in% c(450000, 109000) & n_room %in% c(93, 33)))
```

Check for *NA* percentage in each column:

```{r}
sapply(madrid_idealista, function(x) sum(is.na(x))*100/nrow(madrid_idealista))
```

```{r}
madrid_idealista %>% filter(is.na(neighbourhood)) %>% arrange(id)
```

After a deeper investigation it has been found that those remaining dwellings without neighbourhood information are located in the main streets or avenues which are found in the boundaries of two neighbourhoods. The decision has been to consult manually the location and assign a neighbourhood according to the part of the street where these houses are found. Despite this small detail, neighbourhood itself will not be as important for this study, the interest is mainly in the district, which is the same for the two neighbourhoods. Only dwellings from the three first quarters of the year will be handled because, as explained later, the analysis is only going to be considering this period.

```{r}
missing_n <- data.frame(madrid_idealista %>% 
  filter(id %in% c("A16435294318315958522", "A713741178483582771", "A17974577062442055154", "A7640522779001274113", 
                   "A16293887458042174614", "A2558868114930352875", "A10040860376639148021", "A11024076574113822412", 
                   "A8975421952301567018", "A15391127164769132499", "A7276195934098417167", "A14373658715532580887", 
                   "A12512273801481420708", "A4124869518560963783", "A1047707405959696934", "A14527656687040890533", 
                   "A392049767746391904", "A9152666484316798621", "A10555682128559782831", "A1770016731578591700", 
                   "A2850681705060765887", "A15122532253238651164", "A11875273696714479769", "A2916148036363906874", 
                   "A8323672166361605012", "A14425820252961357301", "A16832610759102053201", "A16175240154493734292", 
                   "A12449866803255143264", "A6566989044715844699", "A4695742828180854443", "A8834840793393699611", 
                   "A2371944396876658222", "A17669068010471096780", "A12275460312855642440", "A13456631820999789083", 
                   "A2415811276745293011", "A12855371497725279279", "A12427375718741803668", "A15415301456868430114", 
                   "A13947664044826473945")) %>% 
  mutate(neighbourhood = c("Legazpi", "Castilla", "Palomeras Bajas", "Piovera", "Pinar del Rey", 
                           "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", 
                           "Casco Histórico de Vallecas", "Adelfas", "Ciudad Jardín", "Nueva España", "Nueva España", 
                           "Ciudad Universitaria", "Valdebebas - Valdefuentes", "Valdefuentes", 
                           "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", 
                           "Alameda de Osuna", "Legazpi", "Estrella", "Valdeacederas", "Castillejos", 
                           "Puerta del Ángel", "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", 
                           "Valdebebas - Valdefuentes", "Arapiles", "Cuatro Caminos", "Buenavista", "Vinateros", 
                           "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", 
                           "Delicias", "Peñagrande", "Valdebebas -Valdefuentes", "Valdebebas - Valdefuentes", 
                           "Valdebebas - Valdefuentes", "Casco Histórico de Vallecas"))) %>% 
  st_drop_geometry()

missing_n
```


```{r}
madrid_idealista <- madrid_idealista %>% 
  filter(!(id %in% c("A16435294318315958522", "A713741178483582771", "A17974577062442055154", "A7640522779001274113", 
                   "A16293887458042174614", "A2558868114930352875", "A10040860376639148021", "A11024076574113822412", 
                   "A8975421952301567018", "A15391127164769132499", "A7276195934098417167", "A14373658715532580887", 
                   "A12512273801481420708", "A4124869518560963783", "A1047707405959696934", "A14527656687040890533", 
                   "A392049767746391904", "A9152666484316798621", "A10555682128559782831", "A1770016731578591700", 
                   "A2850681705060765887", "A15122532253238651164", "A11875273696714479769", "A2916148036363906874", 
                   "A8323672166361605012", "A14425820252961357301", "A16832610759102053201", "A16175240154493734292", 
                   "A12449866803255143264", "A6566989044715844699", "A4695742828180854443", "A8834840793393699611", 
                   "A2371944396876658222", "A17669068010471096780", "A12275460312855642440", "A13456631820999789083", 
                   "A2415811276745293011", "A12855371497725279279", "A12427375718741803668", "A15415301456868430114", 
                   "A13947664044826473945"))) %>% 
  st_drop_geometry()

madrid_idealista <- madrid_idealista %>% bind_rows(missing_n) %>% select(-geometry)
```

Cleaned Idealista's data.

```{r}
madrid_idealista
```

### 3.2. Other external data

#### 3.2.1. Euribor

Scrap of Euribor data:

- Identifying who is scraping with *my user agent*: this step is not compulsory/required, but is highly recommended. Each user should introduce their information regarding profile and device, so that the scraped website is aware of who is doing so and if the user is causing any inconvenience they could be rapidly contacted with no further grievance.

```{r}
# set_config(user_agent("User Agent; Name Surname Surname (occupation or purpose) / email"))
```

- Reading the website into R.

```{r}
link_euribor <- "https://datosmacro.expansion.com/hipotecas/euribor?anio=2018"
html_website <- link_euribor %>% read_html()
html_website
```

- Selecting the desired table.

```{r}
euribor18 <- html_website %>%
  html_table() %>% 
  .[[3]]

euribor18
```

- Handling needed data.

```{r}
euribor18 <- euribor18 %>% 
  select(Fecha, Tipo) %>% 
  rename(month_sp = Fecha, value = Tipo) %>% 
  slice(1:(nrow(euribor18) - 1)) # to remove last row

euribor18 <- euribor18 %>%
  mutate(month_sp = str_replace_all(month_sp, " 2018", ""), # to eliminate "2018" from each row in "Fecha" (date) column
         value = str_replace_all(value, ",", ".")) %>% # to replace commas per dots
  mutate(month = case_when( # to manually translate month names
    month_sp == "Enero" ~ "January",
    month_sp == "Febrero" ~ "February",
    month_sp == "Marzo" ~ "March",
    month_sp == "Abril" ~ "April",
    month_sp == "Mayo" ~ "May",
    month_sp == "Junio" ~ "June",
    month_sp == "Julio" ~ "July",
    month_sp == "Agosto" ~ "August",
    month_sp == "Septiembre" ~ "September",
    month_sp == "Octubre" ~ "October",
    month_sp == "Noviembre" ~ "November",
    month_sp == "Diciembre" ~ "December",
    TRUE ~ NA_character_
  )) %>% 
  select(month, value, -month_sp)

euribor18
```

- Assigning each month its corresponding quarter.

```{r}
euribor18_q <- euribor18 %>% 
  mutate(period = case_when(
    month %in% c("December", "November", "October") ~ "20184",
    month %in% c("September", "August", "July") ~ "20183",
    month %in% c("June", "May", "April") ~ "20182",
    month %in% c("March", "February", "January") ~ "20181"),
         period = as.yearqtr(format(period), "%Y%q"),
         value = str_replace_all(value, "%", ""),
         value = as.numeric(value)
    )
```

- Grouping by quarter and calculating average Euribor per each period.

```{r}
euribor18_q <- euribor18_q %>%
  select(period, value) %>% 
  group_by(period) %>%
  summarise(average_euribor = round(mean(value),4))
  
euribor18_q
```

- Joining with the main dataset.

```{r}
madrid_idealista <- madrid_idealista %>% 
  full_join(euribor18_q, by = "period")
```

#### 3.2.2. Airbnb

Extraction of information regarding number of houses and places present in Airbnb in each corresponding neighbourhood. Then each dwelling will be assigned a number depending the neighbourhood where it is found.

- Reading raw data from .csv file.

```{r}
airbnb <- read_csv("Data/airbnb_data.csv", show_col_types = FALSE)

airbnb
```

- Checking for no duplicates in Airbnb data.

```{r}
sum(duplicated(airbnb$id))
```

This database only contains information up to the third quarter of the year 2018.

```{r}
min(airbnb$found)
max(airbnb$found)
```

```{r}
airbnb <- airbnb %>% 
  select(id, longitude, latitude, found, revised, room_type, capacity) %>% 
  rename(long = longitude, lat = latitude)
```

"Revised" is the last time found the dwelling in the platform. Only observations from 01th January, 2018 will be considered.

```{r}
airbnb <- airbnb %>% 
  filter(revised >= as.Date("2018-01-01")) 
```

- Creating geometry object for each dwelling.

```{r}
airbnb <- st_as_sf(airbnb, coords = c("long", "lat"), crs = 4326)

airbnb
```

- Assigning a neighbourhood to each Airbnb home.

```{r}
airbnb <- st_join(airbnb, Madrid_Polygons, join = st_within) %>% 
  select(-c(LOCATIONID, ZONELEVELID)) %>% 
  rename(neighbourhood = LOCATIONNAME)

sum(is.na(airbnb$neighbourhood))
```

Those 2420 homes that have an *NA* in *neighbourhood* is because they are from the Community of Madrid, but not from the City of Madrid, so there is no interest in working on them. They can be removed from data.

```{r}
airbnb <- airbnb %>% 
  filter(!is.na(neighbourhood))
```

Consult aggregated data by neighbourhood:

```{r}
airbnb %>%
  group_by(neighbourhood) %>%
  summarise(total_capacity = sum(capacity)) %>%
  arrange(desc(total_capacity))
```

- Separating houses in Airbnb by quarters (01 for the first, 02 for the second and 03 for the third).

```{r}
airbnb_q1 <- airbnb %>% 
  filter(found < as.Date("2018-04-01")) 

airbnb_q2 <- airbnb %>% 
  filter(found > as.Date("2018-03-31") & found < as.Date("2018-07-01") |
         revised > as.Date("2018-03-31") & revised < as.Date("2018-07-01"))

airbnb_q3 <- airbnb %>% 
  filter(revised > as.Date("2018-06-30")) 
```

- Adding each dataset its corresponding quarter.

```{r}
airbnb_q1 <- airbnb_q1 %>% 
  mutate(period = as.yearqtr("2018 Q1", format = "%Y Q%q"))

airbnb_q2 <- airbnb_q2 %>% 
  mutate(period = as.yearqtr("2018 Q2", format = "%Y Q%q"))

airbnb_q3 <- airbnb_q3 %>% 
  mutate(period = as.yearqtr("2018 Q3", format = "%Y Q%q"))
```

- Reaching to the final aggregated Airbnb information by quarters.

```{r}
airbnb_ag <- rbind(airbnb_q1, airbnb_q2, airbnb_q3)
```

There are Airbnb dwellings operating in more than one quarter of the year.

```{r}
sum(duplicated(airbnb_ag$id))
```

- Creating the final Airbnb dataset: assigning number of houses and number of places in Airbnb to each neighbourhood according to the quarter of the year.

```{r message = FALSE}
airbnb_final <- airbnb_ag %>%
  group_by(neighbourhood, period) %>%
  summarise(n_houses_airbnb = n(),
            n_places_airbnb = sum(capacity)) %>% 
  st_drop_geometry()

head(airbnb_final, n = 10)
```

- Incorporating sf geometry.

```{r}
madrid_idealista <- st_as_sf(madrid_idealista, coords = c("long", "lat"), crs = 4326) %>% 
  mutate(long = st_coordinates(geometry)[, 1],
         lat = st_coordinates(geometry)[, 2])
```

- Removing Q4 from the main dataset and incorporating Airbnb data into it.

```{r}
madrid_idealista <- madrid_idealista %>% 
  filter(period != "2018 Q4") # because there is no Airbnb information about Q4, we remove this part

madrid_idealista <- madrid_idealista %>%
  full_join(airbnb_final, by = c("neighbourhood", "period")) # to incorporate extracted Airbnb data into the main dataset
```

- Substituting missing values per 0 (neighbourhoods resulting in NA are because they do not have any Airbnb house nearby).

```{r}
madrid_idealista <- madrid_idealista %>%
  mutate(n_houses_airbnb = replace_na(n_houses_airbnb, 0),
         n_places_airbnb = replace_na(n_places_airbnb, 0))
```

#### 3.2.3. Parks and green areas

- Reading raw data from .csv file.

```{r message = FALSE}
parks <- read_delim("Data/parks_green_areas.csv", delim = ";", show_col_types = FALSE) %>% 
  select(c(NOMBRE, LATITUD, LONGITUD)) %>% 
  rename(name = NOMBRE, lat = LATITUD, long = LONGITUD)
```

- Incorporating sf geometry.

```{r}
parks <- st_as_sf(parks, coords = c("long", "lat"), crs = 4326) %>% 
  mutate(long = st_coordinates(geometry)[, 1],
         lat = st_coordinates(geometry)[, 2])
data("Madrid_Polygons")
Madrid_Polygons
```

- Calculating shortest distance to an educational center.

```{r}
madrid_idealista <- madrid_idealista %>%
  mutate(park_proximity_m = st_distance(., parks) %>% apply(1, min),
         park_proximity = park_proximity_m / 1000) %>% 
  select(-park_proximity_m)
```

Aluche, example distance calculation (to Parque Aluche):

```{r}
# A2589730015233487221 is a home in Aluche near a park
# A1827641402258800038 is a home in Aluche a bit further

madrid_idealista %>% select(id, park_proximity) %>% filter(id == "A2589730015233487221" | id == "A1827641402258800038")
```

Opera, example distance calculation (to Jardines de Sabatini):

```{r}
# A12651484524718559564 # is a home in Opera (Palacio) near to Jardines de Sabatini
# A16593819002392137683 # is a home in Opera (Palacio) a bit further
madrid_idealista %>% select(id, park_proximity) %>% filter(id == "A12651484524718559564" | id == "A16593819002392137683")
```

#### 3.2.4. Educational centers

- Reading raw data from .csv file.

```{r message = FALSE}
educational_centers <- read_delim("Data/educational_centers.csv", delim = ";",
                                  show_col_types = FALSE) %>% 
  select(NOMBRE, LONGITUD, LATITUD) %>% 
  rename(name = NOMBRE, long = LONGITUD, lat = LATITUD) %>% 
  drop_na() # drop 5 centers without location
```

- Incorporating sf geometry.

```{r}
educational_centers <- st_as_sf(educational_centers, coords = c("long", "lat"), crs = 4326) %>% 
  mutate(long = st_coordinates(geometry)[, 1],
         lat = st_coordinates(geometry)[, 2])
```

- Calculating shortest distance to an educational center.

```{r}
madrid_idealista <- madrid_idealista %>%
  mutate(educ_center_proximity_m = st_distance(., educational_centers) %>% apply(1, min),
         educ_center_proximity = educ_center_proximity_m / 1000) %>% select(-educ_center_proximity_m)
```

#### 3.2.5. Health centers

- Reading raw data from .csv file.

```{r message = FALSE}
health_centers <- read_delim("Data/health_centers.csv", delim = ";", show_col_types = FALSE) %>% 
  select(NOMBRE, LONGITUD, LATITUD) %>%
  rename(name = NOMBRE, long = LONGITUD, lat = LATITUD)
```

- Removing drug and addictions centers.

```{r message = FALSE}
health_centers <- health_centers %>% 
  filter(!str_detect(name, "CAD|CCAD|CAID"))
```

- Incorporating sf geometry.

```{r}
health_centers <- st_as_sf(health_centers, coords = c("long", "lat"), crs = 4326) %>% 
  mutate(long = st_coordinates(geometry)[, 1],
         lat = st_coordinates(geometry)[, 2])
```

- Calculating shortest distance to an educational center.

```{r}
madrid_idealista <- madrid_idealista %>%
  mutate(health_center_proximity_m = st_distance(., health_centers) %>% apply(1, min),
         health_center_proximity = health_center_proximity_m / 1000) %>%
  select(-health_center_proximity_m)
```

Save the final dataset into a .csv

```{r}
# madrid_idealista_to_save <- as.data.frame(st_drop_geometry(madrid_idealista))
```

```{r}
# write.csv(madrid_idealista_to_save, file = "madrid_idealista.csv", row.names = FALSE, quote = FALSE, fileEncoding = "UTF-8")
```

#### Plot 4: Parks, Educational and Health centers map

[Leaflet](https://rstudio.github.io/leaflet/map_widget.html)

Map creation:

- Empty map.

```{r}
map <- leaflet() %>%
  addTiles()
```

- Adding three different layers: parks, educational_centers and health_centers.

```{r}
# Parks layer
map <- addCircleMarkers(
  data = parks,
  lng = ~long,  # Column containing longitude values
  lat = ~lat,   # Column containing latitude values
  color = "green",
  radius = 5, # Adjust the size of the markers as needed
  stroke = FALSE,
  fillOpacity = 0.7,
  group = "Parks",  # Assign a group name for the layer
  map = map
)

# Educational centers layer
map <- addCircleMarkers(
  data = educational_centers,
  lng = ~long,
  lat = ~lat,
  color = "blue",
  radius = 5,
  stroke = FALSE,
  fillOpacity = 0.7,
  group = "Educational Centers",  # Assign a group name for the layer
  map = map
)

# Health centers layer
map <- addCircleMarkers(
  data = health_centers,
  lng = ~long,
  lat = ~lat,
  color = "red",
  radius = 5,
  stroke = FALSE,
  fillOpacity = 0.7,
  group = "Health Centers",  # Assign a group name for the layer
  map = map
)
```

- Adding a filter control.

```{r}
map <- addLayersControl(
  overlayGroups = c("Parks", "Educational Centers", "Health Centers"),
  options = layersControlOptions(collapsed = FALSE),
  map = map
)
```

- Adding legend.

```{r}
map <- addLegend(
  map = map,
  position = "bottomright",
  colors = c("green", "blue", "red"),
  labels = c("Parks", "Educational Centers", "Health Centers"),
  title = "Facility type"
)
```

- Final map.

```{r}
map
```

### 3.3. Data quality assessment

This section dedicated to exploring characteristics of duplicates in data. 

Dataset with houses repeated more than once (including all rows).

```{r}
exploratory <- idealista18 %>%
  group_by(id) %>%
  filter(n() > 1) %>%
  arrange(id)

exploratory <- exploratory %>%
  group_by(id) %>%
  mutate(across(c(north_orientation, south_orientation, east_orientation,
                  west_orientation), ~max(.)))

exploratory <- exploratory %>%
  mutate(orientation = as_factor(case_when(
  north_orientation == "0" & south_orientation == "0" & east_orientation == "0" & west_orientation == "0" ~ "None",
  north_orientation == "1" & south_orientation == "0" & east_orientation == "0" & west_orientation == "0" ~ "Low", # north
  north_orientation == "0" & south_orientation == "0" & east_orientation == "0" & west_orientation == "1" ~ "Medium", # west
  north_orientation == "0" & south_orientation == "0" & east_orientation == "1" & west_orientation == "0" ~ "Medium", # east
  north_orientation == "0" & south_orientation == "0" & east_orientation == "1" & west_orientation == "1" ~ "Medium", # east-west
  north_orientation == "1" & south_orientation == "0" & east_orientation == "0" & west_orientation == "1" ~ "Medium", # north-west
  north_orientation == "1" & south_orientation == "0" & east_orientation == "1" & west_orientation == "0" ~ "Medium", # north-east
  north_orientation == "1" & south_orientation == "0" & east_orientation == "1" & west_orientation == "1" ~ "Medium", # north-east-west
  north_orientation == "0" & south_orientation == "1" & east_orientation == "0" & west_orientation == "0" ~ "Good", # south
  north_orientation == "0" & south_orientation == "1" & east_orientation == "0" & west_orientation == "1" ~ "Good", # south-west
  north_orientation == "0" & south_orientation == "1" & east_orientation == "1" & west_orientation == "0" ~ "Good", # south-east
  north_orientation == "1" & south_orientation == "1" & east_orientation == "0" & west_orientation == "0" ~ "Good", # south-north
  north_orientation == "0" & south_orientation == "1" & east_orientation == "1" & west_orientation == "1" ~ "Good", # south-east-west
  north_orientation == "1" & south_orientation == "1" & east_orientation == "1" & west_orientation == "0" ~ "Good", # south-north-east
  north_orientation == "1" & south_orientation == "1" & east_orientation == "0" & west_orientation == "1" ~ "Good", # south-north-west
  north_orientation == "1" & south_orientation == "1" & east_orientation == "1" & west_orientation == "1" ~ "Full")),
         period = case_when(period == 201803 ~ "20181",
                            period == 201806 ~ "20182",
                            period == 201809 ~ "20183",
                            period == 201812 ~ "20184"),
         period = as.yearqtr(format(period), "%Y%q")) %>%
  select(-c(north_orientation, south_orientation, east_orientation, west_orientation))

exploratory
```

```{r}
variations <- exploratory %>%
  group_by(id) %>%
  summarise(price_min = min(price_sq_m),
            price_max = max(price_sq_m),
            price_variation = ((price_max - price_min)/price_min)*100)

variations
```

Average variation in the price of a square meter.

```{r}
sum(is.na(variations$price_variation))
mean(variations$price_variation)
```

Orientation: dwellings with no location.

```{r}
exploratory
```

Mean antiquity of buildings with no orientation is more than 55 years. 

```{r}
exploratory_orientation <- exploratory %>%
  group_by(id) %>%
  filter(orientation == "None") %>%
  slice(1) # select the first row that satisfies the condition

mean(exploratory_orientation$year_built)
```

## REFERENCES

CSSLab UC3M (2023). Data Visualization: Principles and Practice. Retrieved April 4, 2023, from https://csslab.uc3m.es/dataviz/

Expansion.com. (January, 2019). Euribor 2018 - Histórico Euribor 2018. Datosmacro. Retrieved April 21, 2023, from https://datosmacro.expansion.com/hipotecas/euribor?anio=2018

Cimentada, J. (2023). Data Harvesting with R. Retrieved March 29, 2023, from https://cimentadaj.github.io/dataharvesting/

Datahippo. (2018, October 1). Datahippo.org. Madrid (Provincia): Datos básicos Airbnb. Retrieved April 17, 2023, from https://datahippo.org/es/region/599230b08a46554edf88466b/

Open Data Portal of Madrid City Council. (2023). Educational centers in Madrid. Retrieved May 3, 2023, from https://datos.madrid.es/sites/v/index.jsp?vgnextoid=f14878a6d4556810VgnVCM1000001d4a900aRCRD&vgnextchannel=374512b9ace9f310VgnVCM100000171f5a0aRCRD

Open Data Portal of Madrid City Council. (2023). Main parks and municipal gardens in Madrid. Retrieved May 3, 2023, from https://datos.madrid.es/portal/site/egob/menuitem.c05c1f754a33a9fbe4b2e4b284f1a5a0/?vgnextoid=dc758935dde13410VgnVCM2000000c205a0aRCRD&vgnextchannel=374512b9ace9f310VgnVCM100000171f5a0aRCRD&vgnextfmt=default

Open Data Portal of Madrid City Council. (2023). Medical care centers in Madrid. Retrieved May 3, 2023, from https://datos.madrid.es/sites/v/index.jsp?vgnextoid=da7437ac37efb410VgnVCM2000000c205a0aRCRD&vgnextchannel=374512b9ace9f310VgnVCM100000171f5a0aRCRD

Rey-Blanco D., Arbues P., Lopez F., Paez A. (2021). idealista18: Idealista 2018 Data Package. R package version 0.1.1. URL: https://paezha.github.io/idealista18/