---
title: "Master's Thesis"
author: "Maria del Mar Escalas Martorell"
date: "`r Sys.Date()`"
output: html_document
description: "This is the first .Rmd document from a set of three. Code provided is complemetary to the main document containing the Master's Thesis in *Forecasting dwelling prices in the City of Madrid: a comprehensive analysis of the Real Estate market*, which can be found in this repository. This markdown contains the code needed for conducting the analysis until point 2. ANALYSIS OF THE SITUATION included."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.align = "center")
```

Load libraries.

```{r message = FALSE, warning = FALSE}
# General libraries:
library(tidyverse)
library(stringr) # regex
library(lubridate) # operations with dates

# Web scraping libraries:
library(xml2)
library(httr)
library(scrapex)
library(rvest) # read_html

# Plots and tables edition:
library(ggplot2)
library(RColorBrewer)
library(sf) # geometry polygons
library(zoo) # date (quarters)

# Times New Roman font import:
library(extrafont)
```

## 1. INTRODUCTION

### 1.1. Motivation of the work
No code

### 1.2. Ojectives
No code

### 1.3. Methodology
No code

### 1.4. Technical requirements and precautions
No code

## 2. ANALYSIS OF THE SITUATION

### 2.1. Prices of dwellings in Spain

#### Plot 1: Housing Price Index (HPI) variation over time (quarters)

Reading raw data from .csv file.

```{r}
hpi <- read_delim("Data/dwelling_price_index_INE.csv", 
                  delim = ";", skip = 1, col_types = "c", 
                  col_names = c("remove1", "Group", "remove2", "remove3", "Period", "HPI")) %>%
  select(-c("remove1", "remove2", "remove3"))

head(hpi)
```

Organizing the dataframe information:

- Substituting *NAs* in *Group* column by *Total National*.
- Substituting *Madrid* in *Group* column by *Community of Madrid*.
- Substituting commas per dots in *HPI* and converting to numeric column type. 

```{r}
hpi <- hpi %>%
  mutate(Group = ifelse(is.na(Group), "Total national", 
                        sub(".*Madrid.*", "Community of Madrid", Group)),
         HPI = as.numeric(sub(",", ".", HPI)),
         Period = str_replace(Period, "T", ""))

head(hpi)
```

Plotting:

- Formatting *Period* information into year quarters.

```{r}
hpi <- hpi %>%
  mutate(Period = as.yearqtr(format(Period), "%Y%q"))
```

- By default, R may not have *Times New Roman* font available for use in plots. However, this font is commonly installed on most operating systems. To ensure its availability in R, the user should consult the following commented chunk.

```{r}
# font_import() # to import all fonts from the device
# loadfonts(device = "win") # to load fonts in the session and use them in plots  
# fonts() # to check the available fonts for use in R
```

- Plot.

```{r fig.height = 5, fig.width = 8.5}
hpi %>% 
  ggplot(aes(y = HPI, x = Period, group = Group, color = Group)) + 
  geom_line(linewidth = 1.3) + 
  labs(x = NULL, y = NULL) +
  theme(legend.title = element_blank(),
        legend.position = c(0.80, 0.20),
        legend.background = element_rect(color = "black", fill = "white"),
        axis.text.x = element_text(angle = 45, vjust = 0.5, color = "black"),
        axis.text.y = element_text(angle = 0, hjust = 1, color = "black"),
        axis.line = element_line(color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white"),
        legend.spacing.y = unit(0.03, "cm"), # to adjust legend box
        legend.key = element_rect(fill = "transparent"),
        text = element_text(family = "Times New Roman", size = 18, color = "black")) +
  scale_color_manual(values = c("darkblue", "orangered")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.5) +
  scale_y_continuous(limits = c(-7.5, 7.5), breaks = seq(-7.5, 7.5, by = 3),
                     labels = function(x) paste0(x, "%")) + # to add "%" symbol to y-labels
  scale_x_yearqtr(format = "%Y", # format = "%YQ%q" to put QX after each label
                  limits = c(min(hpi$Period), max(hpi$Period)),
                  breaks = seq(from = min(hpi$Period), to = max(hpi$Period), by = 1),
                  expand = c(0, 0.5)) +
  geom_vline(xintercept = as.numeric(seq(min(hpi$Period), max(hpi$Period), by = 1)), 
             linetype = "dashed",
             color = "grey")
```

### 2.2. Prices of dwellings in the City of Madrid

#### Data bank from Madrid Council - Statistical information (for plots 2 and 3)

Loading full raw data:

```{r}
madrid_m2 <- read_delim("Data/price_m2_Madrid_district_neighbourhood.csv", 
                        delim = ";", skip = 5, 
                        n_max = 161, # to skip last 4 rows
                        col_names = c("District", "Neighbourhood", "2017", "2018", "2019", "2020", "2021", "2022"),
                        show_col_types = FALSE) %>% 
  mutate(across(starts_with("20"), ~as.numeric(str_replace_all(., "\\.", "")))) # to substitute dots per nothing and convert to numeric

madrid_m2
```

#### Plot 2: Average price per square meter of second-hand dwellings by District in 2022, in euros. Increase in price per square meter: 2017 vs. 2022, in %

```{r}
madrid_districts <- madrid_m2 %>%
  filter(str_detect(Neighbourhood, '^\\d{2}\\.') | Neighbourhood == "Ciudad de Madrid") %>% # to select from the "Neighbourhood" column rows starting by two digits, followed by a dot or rows which are "Ciudad de Madrid"
  rename("remove" = "District", "District" = "Neighbourhood") %>%
  select(-"remove")

madrid_districts
```

```{r}
madrid_districts <- madrid_districts %>%
  mutate(District = gsub("^\\d{2}\\.\\s*", "", District), # to remove two initial digits and point from each row in "District"
         District = ifelse(District == "Ciudad de Madrid", "City of Madrid", District), # to manually translate into English
         Variation = round(((`2022` - `2017`) / `2017`) * 100, 2)) # to create a new column with the increase between first and last year

madrid_districts
```

Plotting:

- Preparing data in long format.

```{r}
madrid_longer <- madrid_districts %>%
  pivot_longer(!District, names_to = c("Year"), values_to = "price per m2")

madrid_longer
```

- Plot.

```{r fig.height = 5, fig.width = 8.5}
madrid_longer %>%
  filter(Year == "2022") %>%
  ggplot(aes(x = reorder(District, `price per m2`), y = `price per m2`, fill = (District == "City of Madrid"))) +
  labs(x = NULL, y = NULL) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_rect(xmin = -Inf, xmax = Inf, ymin = 6900, ymax = 7900, fill = "darkslategray2") +
  geom_text(aes(label = `price per m2`), 
            hjust = 1.1, nudge_x = 0.1, color = "white", family = "Times New Roman", size = 5, fontface = "bold") +
  scale_y_continuous(limits = c(0, 7900), expand = c(0, 0), breaks = NULL) +
  geom_text(data = madrid_longer %>% filter(Year == "Variation"), 
            aes(x = District, y = 7400, label = paste0(`price per m2`, "%")), 
            vjust = 0.35, family = "Times New Roman", size = 5, fontface = "bold") +
  theme_minimal() +
  theme(text = element_text(family = "Times New Roman", size = 20),
        axis.text.y = element_text(vjust = 0.35, color = "black")) +
  scale_fill_manual(values = c("TRUE" = "orangered", "FALSE" = "darkblue"), guide = "none") # to fill "City of Madrid" distinct from distrits
```

#### Plot 3: Average price per sq meter of Districts in Madrid

- Filtering to remove aggregated data from *City of Madrid* and leave only data from individual districts.

```{r}
madrid_map <- madrid_m2 %>% 
  filter(str_detect(Neighbourhood,'^\\d{2}\\.')) %>% # start by two digits followed by a dot
  rename("remove" = "District", "District" = "Neighbourhood") %>% 
  select(-"remove") 

madrid_map
```

- Loading geospatial data to create the map of Madrid and its districts.

```{r}
geodata_districts <- read_sf(dsn = "./Data", # "." means current directory
  layer = "Geodata_Districts")

geodata_districts <- geodata_districts %>% 
  mutate(NOMBRE = case_when(NOMBRE == "San Blas - Canillejas" ~ "San Blas-Canillejas",
                            NOMBRE == "Fuencarral - El Pardo" ~ "Fuencarral-El Pardo",
                            NOMBRE == "Moncloa - Aravaca" ~ "Moncloa-Aravaca",
                            TRUE ~ NOMBRE),
         District = paste0(COD_DIS_TX, ". ", NOMBRE)) %>% select(District)

geodata_districts
```

- Adding to each district its corresponding geometry polygon. 

```{r}
madrid_map <- madrid_map %>% left_join(geodata_districts, by = "District")

madrid_map
```

- Plot. Two digit code of each district and name is placed after code sections for a better understanding.

```{r fig.height = 6, fig.width = 8.5}
madrid_map %>%
ggplot(aes(fill = `2022`, geometry = geometry)) +
  geom_sf(color = "black") +
  scale_fill_gradient(low = "ivory", high = "orangered") +
  labs(x = NULL, y = NULL) +
  theme_void() +
  theme(text = element_text(family="Times New Roman"),
        legend.title = element_blank(),
        legend.key = element_blank(),
        legend.position = c(1.1, 0.8),
        legend.direction = "vertical",
        legend.text = element_text(size = 15),
        plot.margin = margin(0, 0, 0, 0, "pt"))  +
  annotate("text", x = 426800, y = 4472900, label = "Centro", # 01.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 434000, y = 4461000, label = "Arganzuela", # 02.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 456500, y = 4466000, label = "Retiro", # 03.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 457000, y = 4486500, label = "Salamanca", # 04.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 448200, y = 4487500, label = "Chamartín", # 05.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 445500, y = 4490000, label = "Tetuán", # 06.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 425900, y = 4474600, label = "Chamberí", # 07. 
           size = 5, family = "Times New Roman") +
  annotate("text", x = 435000, y = 4487500, label = "Fuencarral-\nEl Pardo", # 08.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 436800, y = 4477500, label = "Moncloa-\nAravaca", # 09.
           size = 4, family = "Times New Roman") +
  annotate("text", x = 433000, y = 4470500, label = "Latina", # 10.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 427500, y = 4463000, label = "Carabanchel", # 11.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 440400, y = 4469200, label = "Usera", # 12.
           size = 3.5, family = "Times New Roman") +
  annotate("text", x = 451000, y = 4460000, label = "Puente de Vallecas", # 13.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 457500, y = 4468000, label = "Moratalaz", # 14.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 459000, y = 4480000, label = "Ciudad \nLineal", # 15.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 446200, y = 4482500, label = "Hortaleza", # 16.
           size = 4, family = "Times New Roman") +
  annotate("text", x = 441000, y = 4466000, label = "Villaverde", # 17.
           size = 3.5, family = "Times New Roman") +
  annotate("text", x = 447500, y = 4467000, label = "Villa de \nVallecas", # 18.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 451000, y = 4472000, label = "Vicálvaro", # 19.
           size = 5, family = "Times New Roman") +
  annotate("text", x = 448800, y = 4476100, label = "San Blas-\nCanillejas", # 20.
           size = 3.5, family = "Times New Roman") +
  annotate("text", x = 451500, y = 4480500, label = "Barajas", # 21.
            size = 5, family = "Times New Roman") +
  geom_segment(aes(x = 429000, y = 4472800, xend = 440400, yend = 4474500), # 01. Centro
               linewidth = 0.5, color = "black") +
  geom_segment(aes(x = 434000, y = 4461500, xend = 440400, yend = 4472500), # 02. Arganzuela
               linewidth = 0.5, color = "black") +
  geom_segment(aes(x = 442500, y = 4473200, xend = 454000, yend = 4466000), # 03. Retiro
               linewidth = 0.5, color = "black") +
  geom_segment(aes(x = 442500, y = 4475800, xend = 454000, yend = 4485600), # 04. Salamanca
               linewidth = 0.5, color = "black") +
  geom_segment(aes(x = 442500, y = 4480000, xend = 445300, yend = 4486600), # 05. Chamartín
               linewidth = 0.5, color = "black") +
  geom_segment(aes(x = 441000, y = 4480000, xend = 444300, yend = 4489000), # 06. Tetuán
               linewidth = 0.5, color = "black") +
  geom_segment(aes(x = 429000, y = 4474500, xend = 440400, yend = 4476200), # 07. Chamberí
               linewidth = 0.5, color = "black") +
  geom_segment(aes(x = 430000, y = 4464000, xend = 436800, yend = 4469200), # 11. Carabanchel
               linewidth = 0.5, color = "black") +
  geom_segment(aes(x = 443000, y = 4469000, xend = 445900, yend = 4460800), # 13. Puente de Vallecas
               linewidth = 0.5, color = "black") +
  geom_segment(aes(x = 445000, y = 4473200, xend = 454000, yend = 4468000), # 14. Moratalaz
               linewidth = 0.5, color = "black") +
  geom_segment(aes(x = 445000, y = 4477000, xend = 456500, yend = 4480000), # 15. Ciudad Lineal
               linewidth = 0.5, color = "black")
```

## REFERENCES

CSSLab UC3M (2023). Data Visualization: Principles and Practice. Retrieved April 4, 2023, from https://csslab.uc3m.es/dataviz/

National Institute for Statistics – Spain. (2023). Índice de Precios de Vivienda [Housing Price Index]. Retrieved March 15, 2023, from https://www.ine.es/jaxiT3/Datos.htm?t=25171

Open Data Portal of Madrid City Council. (2023). Evolution of the price of second-hand housing (€ per square meter) by District and Neighbourhood. Retrieved April 2, 2023, from https://www-s.madrid.es/CSEBD_WBINTER/seleccionSerie.html?numSerie=0504030000202

Open Data Portal of Madrid City Council. (2021). Madrid municipal districts (geospatial data). Retrieved May 15, 2023, from https://www-s.madrid.es/CSEBD_WBINTER/seleccionSerie.html?numSerie=0504030000202
