---
title: "Master's Thesis"
author: "Maria del Mar Escalas Martorell"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.align = "center")
```

Loading libraries.

```{r message = FALSE}
# General libraries:
library(tidyverse)
library(stringr) # regex
library(lubridate) # operations with dates

# Web scraping libraries:
library(xml2)
library(httr)
library(scrapex)
library(rvest) # read_html

# Libraries for editing plots and tables:
library(ggplot2)
library(RColorBrewer)
library(wordcloud)
library(sf) # geometry polygons
library(zoo) # date (quarters)

# Times New Roman font import:
library(extrafont)

# Operations with geodata
library(geosphere)
```

## 0. Introduction

## 1. Objectives

## 2. Methodology

## 3. Presentation of the problem

#### Plot 1: Dwelling Price Index (DPI) variation over time (quarters)

Source: [INE](https://www.ine.es/jaxiT3/Datos.htm?t=25171)

[Dataviz](https://csslab.uc3m.es/dataviz/)

Reading raw data from .csv file.

```{r}
dpi <- read_delim("Data/dwelling_price_index_INE.csv", 
                  delim = ";", 
                  skip = 1,
                  col_types = "c", 
                  col_names = c("remove1", "Group", "remove2", "remove3", "Period", "DPI")) %>% 
  select(-c("remove1", "remove2", "remove3"))

head(dpi)
```

Organizing the dataframe information:

- Substituting *NAs* in *Group* column by *Total National*.
- Substituting *Madrid* in *Group* column by *Community of Madrid*.
- Substituting commas per dots in *DPI* and converting to numeric column type. 

```{r}
dpi <- dpi %>%
  mutate(Group = ifelse(is.na(Group), "Total national", 
                        sub(".*Madrid.*", "Community of Madrid", Group)),
         DPI = as.numeric(sub(",", ".", DPI)),
         Period = str_replace(Period, "T", ""))

head(dpi)
```

Plotting:

- Formatting *Period* information into year quarters.

```{r}
dpi <- dpi %>%
  mutate(Period = as.yearqtr(format(Period), "%Y%q"))
```

- Times New Roman font is not available in R by default. It should be imported.

The decision in this case has been to import all fonts in this device and see that Times New Roman is already installed,
so it can be used now in R.

```{r}
# font_import()
# loadfonts(device = "win")
# fonts() to consult available fonts to use in R
```

- Plot.

```{r fig.height= 5, fig.width= 8}
dpi_plot <- dpi %>% 
  ggplot(aes(y = DPI, x = Period, group = Group, color = Group)) + 
  geom_line(linewidth = 1.3) + 
  labs(x = NULL, y = "DPI") +
  theme(legend.title = element_blank(),
        legend.position = c(0.80, 0.20),
        legend.background = element_rect(color = "black", fill = "white"),
        axis.text.x = element_text(angle = 45, vjust = 0.5, color = "black"),
        axis.text.y = element_text(angle = 0, hjust = 1, color = "black"),
        axis.line = element_line(color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white"),
        legend.spacing.y = unit(0.03, "cm"), # para poner bien la caja de la leyenda
        #legend.margin = margin(t = 0),
        text = element_text(family = "Times New Roman", size = 10, color = "black")) +
  scale_color_manual(values = c("darkblue", "red")) +
  geom_hline(yintercept = 0, 
             linetype = "dashed", 
             color = "black", linewidth = 0.5) +
  scale_y_continuous(limits = c(-7.5, 7.5), breaks = seq(-7.5, 7.5, by = 3)) +
  scale_x_yearqtr(format = "%Y", # format = "%YQ%q" para poner QX detrás de cada label
                  limits = c(min(dpi$Period), max(dpi$Period)),
                  breaks = seq(from = min(dpi$Period), to = max(dpi$Period), by = 1),
                  expand = c(0, 0.5)) +
  geom_vline(xintercept = as.numeric(seq(min(dpi$Period), max(dpi$Period), by = 1)), 
             linetype = "dashed",
             color = "grey")

dpi_plot
```

Saving the plot in the same directory.

```{r}
# ggsave("dpi_plot.png", dpi = 300, width = 10, height = 6, units = "in", dpi_plot)
```

#### Data bank from Madrid Council - Statistical information

Evolution of the average price per square meter by district and neighbourhood (City of Madrid)

Full data:

```{r}
madrid_m2 <- read_delim("Data/price_m2_Madrid_district_neighbourhood.csv", 
                        delim = ";", 
                        skip = 5, 
                        n_max = 161, # to skip last 4 rows
                        col_names = c("District", "Neighbourhood", "2017", "2018", "2019", "2020", "2021", "2022"),
                        show_col_types = FALSE) %>% 
  mutate(across(starts_with("20"), ~as.numeric(str_replace_all(., "\\.", "")))) # substitutes dots per nothing and converts to numeric

madrid_m2
```

#### Plot 2: Madrid city

Evolution of second-hand dwelling prices (€/m2) in Madrid City.

```{r}
madrid_city <- madrid_m2 %>% 
  filter(Neighbourhood == "Ciudad de Madrid") %>% 
  select(-"District")

madrid_city <- madrid_city %>%
  pivot_longer(!Neighbourhood, names_to = "Year", values_to = "price per m2") %>%
  mutate(Year = as.Date(paste0(Year, "-01-01"), format = "%Y-%m-%d"))

madrid_city
```

[Change background color panel](https://stackoverflow.com/questions/33322061/change-background-color-panel-based-on-year-in-ggplot-r)

[Annotations](http://www.sthda.com/english/wiki/ggplot2-texts-add-text-annotations-to-a-graph-in-r-software)

```{r}
ggplot(madrid_city, aes(x = Year, y = `price per m2`, group = Neighbourhood)) +
  ylim(3000, 4000) +
  labs(title = NULL, x = NULL, y = NULL) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  geom_line(linewidth = 0.8) +
  geom_point(shape = 16) +
  geom_text(aes(label = `price per m2`, vjust = -1.5), 
            family = "Times New Roman") +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.x = element_line(color = "black"),
        panel.background = element_rect(fill = "white")) +
  geom_segment(aes(x = Year, xend = Year, y = 3000, yend = `price per m2`), color = "grey") +
  annotate(geom = "text", x = as.Date("2017-06-30"), 
           y = 3100, label = "+17.02%", color = "red", 
           fontface = "bold", alpha = 0.7, 
           family = "Times New Roman", size = 6) +
  annotate(geom = "text", x = as.Date("2018-06-30"), 
           y = 3100, label = "-2.91%", color = "darkgreen", 
           fontface = "bold", alpha = 0.8, 
           family = "Times New Roman", size = 6) +
  annotate(geom = "text", x = as.Date("2019-06-30"), 
           y = 3100, label = "-1.36%", color = "darkgreen", 
           fontface = "bold", alpha = 0.8, 
           family = "Times New Roman", size = 6) +
  annotate(geom = "text", x = as.Date("2020-06-30"),
           y = 3100, label = "+0.49%", color = "red", 
           fontface = "bold", alpha = 0.7, 
           family = "Times New Roman", size = 6) +
  annotate(geom = "text", x = as.Date("2021-06-30"), 
           y = 3100, label = "+6.07%", color = "red", 
           fontface = "bold", alpha = 0.7, 
           family = "Times New Roman", size = 6)
```

#### Plot 3: Average price per sq meter of Districts in Madrid

```{r}
madrid_districts <- madrid_m2 %>% 
  filter(str_detect(Neighbourhood,'^\\d{2}\\.')) %>% # start by two digits followed by a dot
  rename("remove" = "District", "District" = "Neighbourhood") %>% 
  select(-"remove") 

madrid_districts
```

[Geodata 21 districts from Madrid municipality](https://datos.madrid.es/portal/site/egob/menuitem.c05c1f754a33a9fbe4b2e4b284f1a5a0/?vgnextoid=7d6e5eb0d73a7710VgnVCM2000001f4a900aRCRD&vgnextchannel=374512b9ace9f310VgnVCM100000171f5a0aRCRD&vgnextfmt=default)

```{r}
geodata_districts <- read_sf(dsn = "./Data", # "." means current directory
  layer = "Geodata_Districts")

geodata_districts <- geodata_districts %>% 
  mutate(NOMBRE = case_when(NOMBRE == "San Blas - Canillejas" ~ "San Blas-Canillejas",
                            NOMBRE == "Fuencarral - El Pardo" ~ "Fuencarral-El Pardo",
                            NOMBRE == "Moncloa - Aravaca" ~ "Moncloa-Aravaca",
                            TRUE ~ NOMBRE),
         District = paste0(COD_DIS_TX, ". ", NOMBRE)) %>% select(District)

geodata_districts
```

```{r}
madrid_districts <- madrid_districts %>% 
  left_join(geodata_districts, by = "District")

madrid_districts
```

```{r fig.height= 5, fig.width= 5}
ggplot() +
  geom_sf(data = madrid_districts, aes(fill = `2018`, geometry = geometry)) +
  scale_fill_gradient(low = "white", high = "darkred") +
  labs(title = "Price per sq meter in Madrid City by District", subtitle = "Year: 2018",
       x = "Longitude", y = "Latitude") +
  theme_void() +
  theme(text = element_text(family="Times New Roman"),
        legend.title = element_blank(),
        legend.key = element_blank(),
        legend.position = c(1.1, 0.8),
        legend.direction = "vertical",
        plot.margin = margin(0, 0, 0, 0, "pt"))  +
  annotate("text", x = 427500, y = 4479000, label = "Centro", # 1 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 440000, y = 4460000, label = "Arganzuela", # 2 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 460000, y = 4470000, label = "Retiro", # 3 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 460000, y = 4475000, label = "Salamanca", # 4
           size = 4, family = "Times New Roman") +
  annotate("text", x = 448200, y = 4487500, label = "Chamartín", # 5 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 447000, y = 4490000, label = "Tetuán", # 6 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 427500, y = 4477000, label = "Chamberí", # 7 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 435000, y = 4487500, label = "Fuencarral-\nEl Pardo", # 8 OK
           size = 5, family = "Times New Roman") +
  annotate("text", x = 437000, y = 4477500, label = "Moncloa-\nAravaca", # 9 OK
           size = 3.5, family = "Times New Roman") +
  annotate("text", x = 433000, y = 4470000, label = "Latina", # 10 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 427500, y = 4462500, label = "Carabanchel", # 11 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 432500, y = 4460000, label = "Usera", # 12 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 451000, y = 4460000, label = "Puente de Vallecas", # 13 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 460000, y = 4465000, label = "Moratalaz", # 14 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 460000, y = 4480000, label = "Ciudad Lineal", # 15 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 446200, y = 4482500, label = "Hortaleza", # 16 OK
           size = 3.5, family = "Times New Roman") +
  annotate("text", x = 441000, y = 4466000, label = "Villaverde", # 17 OK
           size = 3, family = "Times New Roman") +
  annotate("text", x = 447500, y = 4467000, label = "Villa de \nVallecas", # 18 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 451000, y = 4472500, label = "Vicálvaro", # 19 OK
           size = 4, family = "Times New Roman") +
  annotate("text", x = 449000, y = 4476100, label = "San Blas-\nCanillejas", # 20 OK
           size = 3, family = "Times New Roman") +
  annotate("text", x = 451500, y = 4480000, label = "Barajas", # 21 OK
            size = 4, family = "Times New Roman") 
```

## 4. Data extraction and preprocessing

### 4.1.1. Idealista (2023)

Read personal private credentials (hidden).

User's personal credentials are provided by API's host. It is shown how they should be loaded in R from a
hidden location in the computer. In this case, credentials are stored in a separated folder (not included in the data
available). Instead of using this directory, the person who wishes to reproduce this code should introduce its own
credentials. If the study is not going to be public, no hidden data is needed so, instead, consumer apikey and secret
can be loaded itself.

```{r}
# consumer_apikey <- readLines("../Secret/consumer_apikey.txt")
# consumer_secret <- readLines("../Secret/consumer_secret.txt")
# api_key_secret <- paste0(consumer_apikey, ":", consumer_secret)
```

Using basic authentication.

```{r}
# secret <- base64_enc(api_key_secret)
# 
# req <- httr::POST("https://api.idealista.com/oauth/token",
#        httr::add_headers(
#          "Authorization" = paste("Basic", secret, sep = " "),
#          "Content-Type" = "application/x-www-form-urlencoded;charset=utf-8"
#             ),
#        body = "grant_type=client_credentials&scope=read"
#             ) 
# 
# token <- paste("Bearer", httr::content(req)$access_token)
```

Establishing the desired parameters to search.

```{r}
# site = "https://api.idealista.com/3.5/es/search?"
# loc = "center="
# x = 40.40691
# y = -3.70958
# country = "&country=es"
# maxItems = "&maxItems=50"
# pages = "&numPage=1"
# dist = "&distance="
# distance = 1000
# property = "&propertyType="
# type = "homes"
# operation = "&operation="
# op = "sale"
# chalet = "&chalet=0"
```

Building the url.

```{r}
# url1 <- paste(site, loc, x, ',', y, country, maxItems, pages, dist, distance, 
#              property, type, operation, op, sep = "")
```

Requesting the information through Idealista's API.

```{r}
# res1 <- httr::POST(url1, httr::add_headers("Authorization" = token))
```

Extracting the JSON content.

```{r}
# cont_raw2 <- httr::content(res1)
```


### 4.1.2. Idealista (2018)

Database from Idealista (private online Real Estate portal).

```{r}
# devtools::install_github("paezha/idealista18")
library(idealista18)
```

Data from Madrid 2018.

[idealista18](https://paezha.github.io/idealista18/)

```{r}
data("Madrid_Sale")
class(Madrid_Sale)
```

Initial selection: from 42 variables to 30.

```{r}
idealista18 <- Madrid_Sale %>% 
  select(ASSETID, PERIOD, PRICE, CONSTRUCTEDAREA, UNITPRICE, ROOMNUMBER, 
         BATHNUMBER, HASTERRACE, HASLIFT, HASAIRCONDITIONING,
         ISPARKINGSPACEINCLUDEDINPRICE, HASNORTHORIENTATION, 
         HASSOUTHORIENTATION, HASEASTORIENTATION, HASWESTORIENTATION, 
         HASBOXROOM, HASWARDROBE, HASSWIMMINGPOOL, HASDOORMAN, HASGARDEN, 
         FLATLOCATIONID, CADCONSTRUCTIONYEAR, CADMAXBUILDINGFLOOR, 
         CADDWELLINGCOUNT, DISTANCE_TO_CITY_CENTER, DISTANCE_TO_METRO, 
         LONGITUDE, LATITUDE) %>% 
  rename(id = ASSETID, period = PERIOD, price = PRICE,
         built_area = CONSTRUCTEDAREA, price_sq_m = UNITPRICE,
         n_room = ROOMNUMBER, n_bath = BATHNUMBER, terrace = HASTERRACE,
         lift = HASLIFT, air_cond = HASAIRCONDITIONING,
         parking = ISPARKINGSPACEINCLUDEDINPRICE,
         north_orientation = HASNORTHORIENTATION,
         south_orientation = HASSOUTHORIENTATION, 
         east_orientation = HASEASTORIENTATION,
         west_orientation = HASWESTORIENTATION, boxroom = HASBOXROOM,
         wardrobe = HASWARDROBE, pool = HASSWIMMINGPOOL, 
         doorman = HASDOORMAN, garden = HASGARDEN,
         external = FLATLOCATIONID, year_built = CADCONSTRUCTIONYEAR,
         floors = CADMAXBUILDINGFLOOR, n_dwelling = CADDWELLINGCOUNT,
         km_to_center = DISTANCE_TO_CITY_CENTER,
         metro_proximity = DISTANCE_TO_METRO, long = LONGITUDE, lat = LATITUDE)
```

Issues with raw data:

- There is a difference between unique row number and the total number of rows.

```{r}
nrow(idealista18) - length(unique(idealista18$id))
```

- Many dwellings without orientation.

```{r}
idealista18 %>% filter(north_orientation == 0 & south_orientation == 0 & east_orientation == 0 & west_orientation == 0)
```

Dropping geometry to make operations.

```{r}
idealista18 <- st_drop_geometry(idealista18)
```

Recoding of external. If "yes" = 1, if "no" = 0.

```{r}
idealista18 <- idealista18 %>% 
  mutate(external = case_when(external == 2 ~ 0,
                              external == 1 ~ 1,
                              is.na(external) ~ NA_real_))
```

Operations:

- Max for logical variables (in numeric format)
- Average for numerical (continuous) variables

```{r}
idealista <- idealista18 %>% 
  group_by(id) %>% 
  mutate(across(c(period, price, built_area, n_room, n_bath, 
                  year_built, floors, n_dwelling, 
                  km_to_center, metro_proximity, long, lat), mean),
         across(c(terrace, lift, air_cond, parking, 
                  north_orientation, south_orientation, east_orientation, 
                  west_orientation, boxroom, wardrobe, pool, doorman, 
                  garden, external), ~max(.)),
         across(c(price, built_area, price_sq_m, n_room, n_bath), round))

sapply(idealista, function(x) sum(is.na(x))*100/nrow(idealista)) # % of NAs per row

idealista <- idealista %>% distinct(id, .keep_all = TRUE) %>% drop_na() # drops rows with no information available about 'external'
```

Classifying orientation:

```{r}
idealista <- idealista %>% 
  mutate(orientation = as_factor(case_when(
  north_orientation == "0" & south_orientation == "0" & east_orientation == "0" & west_orientation == "0" ~ "None",
  north_orientation == "1" & south_orientation == "0" & east_orientation == "0" & west_orientation == "0" ~ "Low", # north
  north_orientation == "0" & south_orientation == "0" & east_orientation == "0" & west_orientation == "1" ~ "Medium", # west
  north_orientation == "0" & south_orientation == "0" & east_orientation == "1" & west_orientation == "0" ~ "Medium", # east
  north_orientation == "0" & south_orientation == "0" & east_orientation == "1" & west_orientation == "1" ~ "Medium", # east-west
  north_orientation == "1" & south_orientation == "0" & east_orientation == "0" & west_orientation == "1" ~ "Medium", # north-west
  north_orientation == "1" & south_orientation == "0" & east_orientation == "1" & west_orientation == "0" ~ "Medium", # north-east
  north_orientation == "1" & south_orientation == "0" & east_orientation == "1" & west_orientation == "1" ~ "Medium", # north-east-west
  north_orientation == "0" & south_orientation == "1" & east_orientation == "0" & west_orientation == "0" ~ "Good", # south
  north_orientation == "0" & south_orientation == "1" & east_orientation == "0" & west_orientation == "1" ~ "Good", # south-west
  north_orientation == "0" & south_orientation == "1" & east_orientation == "1" & west_orientation == "0" ~ "Good", # south-east
  north_orientation == "1" & south_orientation == "1" & east_orientation == "0" & west_orientation == "0" ~ "Good", # south-north
  north_orientation == "0" & south_orientation == "1" & east_orientation == "1" & west_orientation == "1" ~ "Good", # south-east-west
  north_orientation == "1" & south_orientation == "1" & east_orientation == "1" & west_orientation == "0" ~ "Good", # south-north-east
  north_orientation == "1" & south_orientation == "1" & east_orientation == "0" & west_orientation == "1" ~ "Good", # south-north-west
  north_orientation == "1" & south_orientation == "1" & east_orientation == "1" & west_orientation == "1" ~ "Full")),
         period = case_when(period == 201803 ~ "20181",
                            period == 201806 ~ "20182",
                            period == 201809 ~ "20183",
                            period == 201812 ~ "20184"),
         period = as.yearqtr(format(period), "%Y%q")) %>% 
  select(-c(north_orientation, south_orientation, east_orientation, west_orientation))
```

Reconverting the dataframe into a geometry object.

```{r}
idealista <- st_as_sf(idealista, coords = c("long", "lat"), crs = 4326)
```

Joining each dwelling with its corresponding neighbourhood. 

```{r}
madrid_idealista <- st_join(idealista, Madrid_Polygons, join = st_within)

madrid_idealista <- madrid_idealista %>% 
  select(-c(LOCATIONID, ZONELEVELID)) %>% 
  rename(neighbourhood = LOCATIONNAME)
```

To have in the same dataset geometry and separated longitude and latitude information.

```{r}
madrid_idealista <- madrid_idealista %>%
  mutate(long = st_coordinates(geometry)[, 1],
         lat = st_coordinates(geometry)[, 2])
```

Identification of observations that contain errors:

- A10991866817670907438: price = 450000, n_rooms = 93 (wrong number of rooms)
- A8866129769871540210: price = 109000, n_rooms = 33 (wrong number of rooms) 

```{r}
madrid_idealista %>% filter(n_room > 25)
```

- A7501251371917084834: long = - 2.753303, lat = 36.75639 (house located in Almería)

```{r}
madrid_idealista %>% filter(long > -3.50 & lat < 40.00)
```

Removing wrong observations.

```{r}
madrid_idealista <- madrid_idealista %>%
  filter(!(long > -3.50 & lat < 40.00)) %>%
  filter(!(price %in% c(450000, 109000) & n_room %in% c(93, 33)))
```

Check for NA percentage in each variable.

```{r}
sapply(madrid_idealista, function(x) sum(is.na(x))*100/nrow(madrid_idealista))
```

```{r}
madrid_idealista %>% filter(is.na(neighbourhood)) %>% arrange(id)
```

After a deeper investigation it has been found that those remaining dwellings without neighbourhood information are
located in the main streets or avenues which are located in the boundaries of two neighbourhoods.
The best solution found has been to consult manually the location and assign a neighbourhood according to the part of
the street where these houses are found. Despite this small detail, neighbourhood itself will not be as important for
this study, the interest is mainly in the district, which is the same for the two neighbourhoods.
Only dwellings from the three first quarters of the year will be handled because, as explained later, the analysis is
only going to be considering this period.

```{r}
missing_n <- data.frame(madrid_idealista %>% 
  filter(id %in% c("A16435294318315958522", "A713741178483582771", "A17974577062442055154", "A7640522779001274113", 
                   "A16293887458042174614", "A2558868114930352875", "A10040860376639148021", "A11024076574113822412", 
                   "A8975421952301567018", "A15391127164769132499", "A7276195934098417167", "A14373658715532580887", 
                   "A12512273801481420708", "A4124869518560963783", "A1047707405959696934", "A14527656687040890533", 
                   "A392049767746391904", "A9152666484316798621", "A10555682128559782831", "A1770016731578591700", 
                   "A2850681705060765887", "A15122532253238651164", "A11875273696714479769", "A2916148036363906874", 
                   "A8323672166361605012", "A14425820252961357301", "A16832610759102053201", "A16175240154493734292", 
                   "A12449866803255143264", "A6566989044715844699", "A4695742828180854443", "A8834840793393699611", 
                   "A2371944396876658222", "A17669068010471096780", "A12275460312855642440", "A13456631820999789083", 
                   "A2415811276745293011", "A12855371497725279279", "A12427375718741803668", "A15415301456868430114", 
                   "A13947664044826473945")) %>% 
  mutate(neighbourhood = c("Legazpi", "Castilla", "Palomeras Bajas", "Piovera", "Pinar del Rey", 
                           "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", 
                           "Casco Histórico de Vallecas", "Adelfas", "Ciudad Jardín", "Nueva España", "Nueva España", 
                           "Ciudad Universitaria", "Valdebebas - Valdefuentes", "Valdefuentes", 
                           "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", 
                           "Alameda de Osuna", "Legazpi", "Estrella", "Valdeacederas", "Castillejos", 
                           "Puerta del Ángel", "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", 
                           "Valdebebas - Valdefuentes", "Arapiles", "Cuatro Caminos", "Buenavista", "Vinateros", 
                           "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", "Valdebebas - Valdefuentes", 
                           "Delicias", "Peñagrande", "Valdebebas -Valdefuentes", "Valdebebas - Valdefuentes", 
                           "Valdebebas - Valdefuentes", "Casco Histórico de Vallecas"))) %>% 
  st_drop_geometry()

missing_n

madrid_idealista <- madrid_idealista %>% 
  filter(!(id %in% c("A16435294318315958522", "A713741178483582771", "A17974577062442055154", "A7640522779001274113", 
                   "A16293887458042174614", "A2558868114930352875", "A10040860376639148021", "A11024076574113822412", 
                   "A8975421952301567018", "A15391127164769132499", "A7276195934098417167", "A14373658715532580887", 
                   "A12512273801481420708", "A4124869518560963783", "A1047707405959696934", "A14527656687040890533", 
                   "A392049767746391904", "A9152666484316798621", "A10555682128559782831", "A1770016731578591700", 
                   "A2850681705060765887", "A15122532253238651164", "A11875273696714479769", "A2916148036363906874", 
                   "A8323672166361605012", "A14425820252961357301", "A16832610759102053201", "A16175240154493734292", 
                   "A12449866803255143264", "A6566989044715844699", "A4695742828180854443", "A8834840793393699611", 
                   "A2371944396876658222", "A17669068010471096780", "A12275460312855642440", "A13456631820999789083", 
                   "A2415811276745293011", "A12855371497725279279", "A12427375718741803668", "A15415301456868430114", 
                   "A13947664044826473945"))) %>% 
  st_drop_geometry()

madrid_idealista <- madrid_idealista %>% bind_rows(missing_n) %>% select(-geometry)
```

Cleaned idealista's data.

```{r}
madrid_idealista
```

### 4.2.1. Euribor

Identification of my user agent.
This step is not compulsory/required to scrap, but is highly recommended. Each user should introduce their information
regarding profile and device, so that the web being scraped is aware of who is doing so and if the user is causing any
inconvenience they could be rapidly contacted with no further grievance.

```{r}
# set_config(user_agent("User Agent; Name Surname Surname (occupation or purpose) / email"))
```

Scraping Euribor information.

[Expansión/Datosmacro.com](https://datosmacro.expansion.com/hipotecas/euribor?anio=2018)

Reading the website into R.

```{r}
link_euribor <- "https://datosmacro.expansion.com/hipotecas/euribor?anio=2018"
html_website <- link_euribor %>% read_html()
html_website
```

Selecting the desired table.

```{r}
euribor18 <- html_website %>%
  html_table() %>% 
  .[[3]]

euribor18
```

Handling needed data.

```{r}
euribor18 <- euribor18 %>% 
  select(Fecha, Tipo) %>% 
  rename(month_sp = Fecha, value = Tipo) %>% 
  slice(1:(nrow(euribor18) - 1)) # removing last row

euribor18 <- euribor18 %>%
  mutate(month_sp = str_replace_all(month_sp, " 2018", ""), # eliminating "2018" from each row
         value = str_replace_all(value, ",", ".")) %>% 
  mutate(month = case_when(
    month_sp == "Enero" ~ "January",
    month_sp == "Febrero" ~ "February",
    month_sp == "Marzo" ~ "March",
    month_sp == "Abril" ~ "April",
    month_sp == "Mayo" ~ "May",
    month_sp == "Junio" ~ "June",
    month_sp == "Julio" ~ "July",
    month_sp == "Agosto" ~ "August",
    month_sp == "Septiembre" ~ "September",
    month_sp == "Octubre" ~ "October",
    month_sp == "Noviembre" ~ "November",
    month_sp == "Diciembre" ~ "December",
    TRUE ~ NA_character_
  )) %>% 
  select(month, value, -month_sp)
```

```{r}
euribor18_q <- euribor18 %>% 
  mutate(period = case_when(
    month %in% c("December", "November", "October") ~ "20184",
    month %in% c("September", "August", "July") ~ "20183",
    month %in% c("June", "May", "April") ~ "20182",
    month %in% c("March", "February", "January") ~ "20181"),
         period = as.yearqtr(format(period), "%Y%q"),
         value = str_replace_all(value, "%", ""),
         value = as.numeric(value)
    )

euribor18_q <- euribor18_q %>%
  select(period, value) %>% 
  group_by(period) %>%
  summarise(average_euribor = round(mean(value),4))
  
euribor18_q
```

Saving **euribor_q** table.

```{r}
# write.csv(euribor18_q, file = "euribor_q.csv", row.names = FALSE, quote = FALSE, fileEncoding = "UTF-8")
```

```{r}
madrid_idealista <- madrid_idealista %>% 
  full_join(euribor18_q, by = "period")
```

### 4.2.2. Airbnb

[Datahippo Airbnb](https://datahippo.org/media/regions/4c7e35e4-46d3-4a82-baae-1a16d7d0ea48/599230508a46554e81f2887e_airbnb.csv)

```{r}
airbnb <- read_csv("Data/airbnb_data.csv", show_col_types = FALSE)

airbnb
```

Check for no duplicates in Airbnb data.

```{r}
sum(duplicated(airbnb$id))
```

This database only cointains information up to the third quarter of the year 2018.

```{r}
min(airbnb$found)
max(airbnb$found)
```

```{r}
airbnb <- airbnb %>% 
  select(id, longitude, latitude, found, revised, room_type, capacity) %>% 
  rename(long = longitude, lat = latitude)
```

"Revised" is the last time found the dwelling in the platform. We are only going to consider those from 01th of January of 2018.

```{r}
airbnb <- airbnb %>% 
  filter(revised >= as.Date("2018-01-01")) 
```

Creating the geometry for each dwelling.

```{r}
airbnb <- st_as_sf(airbnb, coords = c("long", "lat"), crs = 4326)

airbnb
```

```{r}
airbnb <- st_join(airbnb, Madrid_Polygons, join = st_within) %>% 
  select(-c(LOCATIONID, ZONELEVELID)) %>% 
  rename(neighbourhood = LOCATIONNAME)
```

Those homes that have an NA in their LOCATION is because they are from the Community of Madrid, but not from the City of Madrid, so there is no interest in working on them. They can be removed from the data.

```{r}
airbnb <- airbnb %>% 
  filter(!is.na(neighbourhood))
```

Consult aggregated data by neighbourhood.

```{r}
airbnb %>%
  group_by(neighbourhood) %>%
  summarise(total_capacity = sum(capacity)) %>%
  arrange(desc(total_capacity))
```

Separating houses in Airbnb by quarters (01 for the first, 02 for the second and 03 for the third).

```{r}
airbnb_q1 <- airbnb %>% 
  filter(found < as.Date("2018-04-01")) 

airbnb_q2 <- airbnb %>% 
  filter(found < as.Date("2018-07-01") & revised > as.Date("2018-03-30") & revised < as.Date("2018-07-01")) 

airbnb_q3 <- airbnb %>% 
  filter(revised > as.Date("2018-06-30")) 
```

Adding each dataset its corresponding quarter.

```{r}
airbnb_q1 <- airbnb_q1 %>% 
  mutate(period = as.yearqtr("2018 Q1", format = "%Y Q%q"))

airbnb_q2 <- airbnb_q2 %>% 
  mutate(period = as.yearqtr("2018 Q2", format = "%Y Q%q"))

airbnb_q3 <- airbnb_q3 %>% 
  mutate(period = as.yearqtr("2018 Q3", format = "%Y Q%q"))
```

Reaching to the final aggregated Airbnb information by quarters.

```{r}
airbnb_ag <- rbind(airbnb_q1, airbnb_q2, airbnb_q3)
```

There are Airbnb dwellings operating in more than one quarter of the year.

```{r}
sum(duplicated(airbnb_ag$id))
```

```{r message = FALSE}
airbnb_final <- airbnb_ag %>%
  group_by(neighbourhood, period) %>%
  summarise(n_houses_airbnb = n(),
            n_places_airbnb = sum(capacity)) %>% 
  st_drop_geometry()

airbnb_final
```

Saving **airbnb_final** table.

```{r}
# write.csv(airbnb_final, file = "airbnb_final.csv", row.names = FALSE, quote = FALSE, fileEncoding = "UTF-8")
```

To have in the same dataset geometry and separated longitude and latitude information.

```{r}
madrid_idealista <- st_as_sf(madrid_idealista, coords = c("long", "lat"), crs = 4326) %>% 
  mutate(long = st_coordinates(geometry)[, 1],
         lat = st_coordinates(geometry)[, 2])
```

Neighbourhoods resulting in NA are because they do not have any Airbnb house surrounding.

```{r}
madrid_idealista <- madrid_idealista %>% 
  filter(period != "2018 Q4")

madrid_idealista <- madrid_idealista %>% 
  # st_drop_geometry() %>% 
  full_join(airbnb_final, by = c("neighbourhood", "period"))

madrid_idealista <- madrid_idealista %>%
  mutate(n_houses_airbnb = replace_na(n_houses_airbnb, 0),
         n_places_airbnb = replace_na(n_places_airbnb, 0))
```

### 4.2.3. Parks and green areas

```{r message = FALSE}
parks <- read_delim("Data/parks_green_areas.csv", delim = ";", show_col_types = FALSE) %>% 
  select(c(NOMBRE, LATITUD, LONGITUD)) %>% 
  rename(name = NOMBRE, lat = LATITUD, long = LONGITUD)
```

```{r}
parks <- st_as_sf(parks, coords = c("long", "lat"), crs = 4326) %>% 
  mutate(long = st_coordinates(geometry)[, 1],
         lat = st_coordinates(geometry)[, 2])
data("Madrid_Polygons")
Madrid_Polygons
```

```{r}
madrid_idealista <- madrid_idealista %>%
  mutate(park_proximity = st_distance(., parks) %>% apply(1, min),
         park_proximity_km = park_proximity / 1000) %>% select(-park_proximity)
```

Aluche example distance calculation (to Parque Aluche)

```{r}
# A2589730015233487221 Aluche ~= coord park
# A1827641402258800038 Aluche a bit further

madrid_idealista %>% select(id, park_proximity_km) %>% filter(id == "A2589730015233487221" | id == "A1827641402258800038")
```

Opera example distance calculation (to Jardines de Sabatini)

```{r}
# A12651484524718559564 Opera (Palacio) near to Jardines de Sabatini
# A16593819002392137683 Opera (Palacio) a bit further
madrid_idealista %>% select(id, park_proximity_km) %>% filter(id == "A12651484524718559564" | id == "A16593819002392137683")
```

### 4.2.4. Educational centers

[Educational centers](https://datos.madrid.es/sites/v/index.jsp?vgnextoid=f14878a6d4556810VgnVCM1000001d4a900aRCRD&vgnextchannel=374512b9ace9f310VgnVCM100000171f5a0aRCRD)

```{r message = FALSE}
educational_centers <- read_delim("Data/educational_centers.csv", delim = ";",
                                  show_col_types = FALSE) %>% 
  select(NOMBRE, LONGITUD, LATITUD) %>% 
  rename(name = NOMBRE, long = LONGITUD, lat = LATITUD) %>% 
  drop_na() # drop 5 centers without location
```

Converting to an sf object.

```{r}
educational_centers <- st_as_sf(educational_centers, coords = c("long", "lat"), crs = 4326) %>% 
  mutate(long = st_coordinates(geometry)[, 1],
         lat = st_coordinates(geometry)[, 2])
```

```{r}
madrid_idealista <- madrid_idealista %>%
  mutate(educ_center_proximity = st_distance(., educational_centers) %>% apply(1, min),
         educ_center_proximity_km = educ_center_proximity / 1000) %>% select(-educ_center_proximity)
```

### 4.2.5. Health centers

[Health centers](https://datos.madrid.es/sites/v/index.jsp?vgnextoid=da7437ac37efb410VgnVCM2000000c205a0aRCRD&vgnextchannel=374512b9ace9f310VgnVCM100000171f5a0aRCRD)

```{r message = FALSE}
health_centers <- read_delim("Data/health_centers.csv", delim = ";", show_col_types = FALSE) %>% 
  select(NOMBRE, LONGITUD, LATITUD) %>%
  rename(name = NOMBRE, long = LONGITUD, lat = LATITUD)
```

Remove drug and addictions centers.

```{r message = FALSE}
health_centers <- health_centers %>% 
  filter(!str_detect(name, "CAD|CCAD|CAID"))
```

```{r}
health_centers <- st_as_sf(health_centers, coords = c("long", "lat"), crs = 4326) %>% 
  mutate(long = st_coordinates(geometry)[, 1],
         lat = st_coordinates(geometry)[, 2])
```

```{r}
madrid_idealista <- madrid_idealista %>%
  mutate(health_center_proximity = st_distance(., health_centers) %>% apply(1, min),
         health_center_proximity_km = health_center_proximity / 1000) %>%
  select(-health_center_proximity)
```

Saving the final dataset into a .csv

```{r}
# madrid_idealista_to_save <- as.data.frame(st_drop_geometry(madrid_idealista))
```

```{r}
# write.csv(madrid_idealista_to_save, file = "madrid_idealista.csv", row.names = FALSE, quote = FALSE, fileEncoding = "UTF-8")
```

## 5. Exploratory data analysis

Dataset with houses repeated more than once (including all rows).

```{r}
exploratory <- idealista18 %>% 
  group_by(id) %>%
  filter(n() > 1) %>% 
  arrange(id)

exploratory <- exploratory %>%
  group_by(id) %>% 
  mutate(across(c(north_orientation, south_orientation, east_orientation, 
                  west_orientation), ~max(.)))

exploratory <- exploratory %>% 
  mutate(orientation = as_factor(case_when(
  north_orientation == "0" & south_orientation == "0" & east_orientation == "0" & west_orientation == "0" ~ "None",
  north_orientation == "1" & south_orientation == "0" & east_orientation == "0" & west_orientation == "0" ~ "Low", # north
  north_orientation == "0" & south_orientation == "0" & east_orientation == "0" & west_orientation == "1" ~ "Medium", # west
  north_orientation == "0" & south_orientation == "0" & east_orientation == "1" & west_orientation == "0" ~ "Medium", # east
  north_orientation == "0" & south_orientation == "0" & east_orientation == "1" & west_orientation == "1" ~ "Medium", # east-west
  north_orientation == "1" & south_orientation == "0" & east_orientation == "0" & west_orientation == "1" ~ "Medium", # north-west
  north_orientation == "1" & south_orientation == "0" & east_orientation == "1" & west_orientation == "0" ~ "Medium", # north-east
  north_orientation == "1" & south_orientation == "0" & east_orientation == "1" & west_orientation == "1" ~ "Medium", # north-east-west
  north_orientation == "0" & south_orientation == "1" & east_orientation == "0" & west_orientation == "0" ~ "Good", # south
  north_orientation == "0" & south_orientation == "1" & east_orientation == "0" & west_orientation == "1" ~ "Good", # south-west
  north_orientation == "0" & south_orientation == "1" & east_orientation == "1" & west_orientation == "0" ~ "Good", # south-east
  north_orientation == "1" & south_orientation == "1" & east_orientation == "0" & west_orientation == "0" ~ "Good", # south-north
  north_orientation == "0" & south_orientation == "1" & east_orientation == "1" & west_orientation == "1" ~ "Good", # south-east-west
  north_orientation == "1" & south_orientation == "1" & east_orientation == "1" & west_orientation == "0" ~ "Good", # south-north-east
  north_orientation == "1" & south_orientation == "1" & east_orientation == "0" & west_orientation == "1" ~ "Good", # south-north-west
  north_orientation == "1" & south_orientation == "1" & east_orientation == "1" & west_orientation == "1" ~ "Full")),
         period = case_when(period == 201803 ~ "20181",
                            period == 201806 ~ "20182",
                            period == 201809 ~ "20183",
                            period == 201812 ~ "20184"),
         period = as.yearqtr(format(period), "%Y%q")) %>% 
  select(-c(north_orientation, south_orientation, east_orientation, west_orientation))

exploratory
```

```{r}
variations <- exploratory %>% 
  group_by(id) %>% 
  summarise(price_min = min(price_sq_m),
            price_max = max(price_sq_m),
            price_variation = ((price_max - price_min)/price_min)*100)

variations
```

Average variation in the price of a square meter.

```{r}
sum(is.na(variations$price_variation))
mean(variations$price_variation)
```

Orientation

```{r}
exploratory
```

```{r}
exploratory_orientation <- exploratory %>% 
  group_by(id) %>% 
  filter(orientation == "None") %>% 
  slice(1)
 
mean(exploratory_orientation$year_built)
```



